---
title: "ESM 244 Lab 1"
author: "Roni Shen"
date: "1/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(sf)
library(tmap)
```

### Read in SF trees data

```{r}
sf_trees <- read_csv(here("data", "sf_trees", "sf_trees.csv"))
```

### Basic wrangling reminders

Refresh skills for data wrangling and summary statistics using functions in the 'dplyr' package.

Find the top 5 highest observations of trees by legal_status, do some wrangling, make a graph.

```{r}
top_5_status <- sf_trees %>% 
  count(legal_status) %>% 
    # acts as group_by() + n() + summarise() functions
    # can include multiple variables to group by
  drop_na(legal_status) %>% 
  rename(tree_count = n) %>% 
    # new name goes on left, old name on right
  relocate(tree_count) %>% 
    # move columns to various positions
  slice_max(tree_count, n = 5)
    # arranges by count and picks top 5 (better than arrange/head)
```

Make a graph of those top 5 observations by legal status.

```{r}
ggplot(data = top_5_status, aes(x = fct_reorder(legal_status, tree_count), y = tree_count))+
    # fct_reorder() changes to factor and reorders by variable
  geom_col()+
  labs(x = "Legal Status", y = "Tree Count")+
  coord_flip()+
    # flip axes
  theme_minimal()
```

### A few more data wrangling refresher examples

Only want to keep observations (rows) for Blackwood Acacia trees.

```{r}
blackwood_acacia <- sf_trees %>% 
  filter(str_detect(species, "Blackwood Acacia")) %>% 
    # detect any string within specified variable
  select(legal_status, date, latitude, longitude)
    # select columns to retain

ggplot(data = blackwood_acacia, aes(x = longitude, y = latitude))+
  geom_point()
```

### tidyr::separate() and unite() functions

Useful for combining or separating columns.

Example: tidyr::separate()

```{r}
sf_trees_sep <- sf_trees %>% 
  separate(species, into = c("spp_scientific", "spp_common"), sep = "::")
```

Example: tidyr::unite()

```{r}
sf_trees_unite <- sf_trees %>% 
  unite("id_status", tree_id:legal_status, sep = "_cool!_")
```

